---
output:
  pdf_document: default
  html_document: default
  word_document: default
---
# Assignment: ASSIGNMENT 7
# Name: Parker, Breanna
# Date: 2023-1-30

## Set the working directory to the root of your DSC 520 directory
```{r}
#setwd("C:/Users/brean/OneDrive/Desktop/NucampFolder/projects/dsc520-1")

```


## Load the `data/r4ds/heights.csv` to
```{r}
heights_df <- read.csv("C:/Users/brean/OneDrive/Desktop/NucampFolder/projects/dsc520-1/data/r4ds/heights.csv")

```



# Fit a linear model
```{r}
earn_lm <-  lm(earn ~ height + age + ed + race + sex, data=heights_df)

```



# View the summary of your model

```{r}

summary(earn_lm)

```


```{r}
predicted_df <- data.frame(
  earn = predict(earn_lm, heights_df),
  ed=heights_df, race=heights_df$race, height=heights_df$height,
  age=heights_df$age, sex=heights_df$sex
  )

```




## Compute deviation (i.e. residuals)

```{r}
mean_earn <- mean(heights_df$earn)
mean_earn

```



## Corrected Sum of Squares Total

```{r}
sst <- sum((mean_earn - heights_df$earn)^2)
sst

```


## Corrected Sum of Squares for Model

```{r}
ssm <- sum((mean_earn - predicted_df$earn)^2)
ssm

```


## Residuals

```{r}

residuals <- heights_df$earn - predicted_df$earn

```



## Sum of Squares for Error

```{r}
sse <- sum(residuals^2)

```



## R Squared

```{r}
r_squared <- ssm/sst
r_squared

```


## Number of observations
```{r}
n <- length(heights_df$age)
n

```


## Number of regression paramaters

```{r}
p <- 8

```



## Corrected Degrees of Freedom for Model

```{r}
dfm <- p-1
dfm

```



## Degrees of Freedom for Error

```{r}
dfe <- n - p
dfe

```



## Corrected Degrees of Freedom Total:   DFT = n - 1

```{r}
dft <- n - 1
dft

```



## Mean of Squares for Model:   MSM = SSM / DFM

```{r}
msm <- ssm/dfm
msm

```



## Mean of Squares for Error:   MSE = SSE / DFE
```{r}
mse <- sse/dfe
mse

```


## Mean of Squares Total:   MST = SST / DFT

```{r}
mst <- sst/dft
mst

```



## F Statistic

```{r}
f_score <- msm/mse
f_score

```




## Adjusted R Squared R2 = 1 - (1 - R2)(n - 1) / (n - p)

```{r}
adjusted_r_squared <-  1 - (1- r_squared)*(n - 1)/ (n - p)
adjusted_r_squared

```

```{r}
names(summary(earn_lm))
summary(earn_lm)$r.squared
summary(earn_lm)$adj.r.squared
summary(earn_lm)$fstatistic

```


