---
output:
  word_document: default
  html_document: default
---
# Assignment: ASSIGNMENT 6
# Name: Parker, Breanna
# Date: 2023-01-30

## Set the working directory to the root of your DSC 520 directory
```{r}
setwd("C:/Users/brean/OneDrive/Desktop/NucampFolder/projects/dsc520-1")

```



## Load the `data/r4ds/heights.csv` to

```{r}

heights_df <- read.csv("C:/Users/brean/OneDrive/Desktop/NucampFolder/projects/dsc520-1/data/r4ds/heights.csv")

```




## Load the ggplot2 library

```{r}

library(ggplot2)

```



## Fit a linear model using the `age` variable as the predictor and `earn` as the outcome
```{r}

age_lm <-  lm(heights_df$age~heights_df$earn)

```




## View the summary of your model using `summary()`

```{r}
summary(age_lm)


```



## Creating predictions using `predict()`

```{r}

age_predict_df <- data.frame(earn = predict(age_lm, heights_df), age=heights_df$age)


```




## Plot the predictions against the original data

```{r}
ggplot(data = heights_df, aes(y = earn, x = age)) +
  geom_point(color='blue') +
  geom_line(color='red',data = age_predict_df, aes(y = earn, x = age))

```




```{r}
mean_earn <- mean(heights_df$earn)
mean_earn

```


## Corrected Sum of Squares Total

```{r}
sst <- sum((mean_earn - heights_df$earn)^2)
sst

```


## Corrected Sum of Squares for Model
```{r}
ssm <- sum((mean_earn - age_predict_df$earn)^2)
ssm

```


## Residuals
```{r}
residuals <- heights_df$earn - age_predict_df$earn

```


## Sum of Squares for Error
```{r}
sse <- sum(residuals^2)
sse

```



## R Squared R^2 = SSM\SST

```{r}
r_squared <- ssm/sst
r_squared

```



## Number of observations

```{r}
n <- length(heights_df$age)
n

```


## Number of regression parameters
```{r}
p <- 2

```


## Corrected Degrees of Freedom for Model (p-1)

```{r}

dfm <- p-1
dfm

```


## Degrees of Freedom for Error (n-p)
```{r}

dfe <- n - p
dfe

```


## Corrected Degrees of Freedom Total:   DFT = n - 1

```{r}
dft <- n - 1
dft

```



## Mean of Squares for Model:   MSM = SSM / DFM

```{r}
msm <- ssm/dfm
msm

```



## Mean of Squares for Error:   MSE = SSE / DFE

```{r}

mse <- sse/dfe
mse

```



## Mean of Squares Total:   MST = SST / DFT

```{r}

mst <- sst/dft
mst

```



## F Statistic F = MSM/MSE
```{r}

f_score <- msm/mse
f_score

```




## Adjusted R Squared R2 = 1 - (1 - R2)(n - 1) / (n - p)
```{r}

adjusted_r_squared <- 1 - (1- r_squared)*(n - 1)/ (n - p)
adjusted_r_squared

```



## Calculate the p-value from the F distribution

```{r}

p_value <- pf(f_score, dfm, dft, lower.tail=F)
p_value

```


```{r}
names(summary(age_lm))
summary(age_lm)$r.squared
summary(age_lm)$adj.r.squared
summary(age_lm)$fstatistic
```

